Building Metagenome Viral Set and Visualization Template

### 2023-05-29 EnvironmentalData/identifying_viral_contigs_environmental.Rmd
Updated environmental scripts

---
title: "Building Environmental Viral Set and Visualization"
author: Bridget Hegarty, James Riddell
date: 10-26-2022
output: html_notebook
---
This Rmarkdown file takes the output from "building_env_data_viral_features_set.Rmd" and creates a set of viral contigs and visualizations.

@@

TO DO:
- add kaiju and DVF

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

```{r}
# dataset name for file organization and outputs for metrics file
dataset_name <- "drinking_water" #replace with the dataset_name from "building_env_data_viral_features_set.Rmd"
#name from building_env_data_viral_features_set.Rmd
viruses_filename <- 'merged_viral_feature_outputs_nluk.csv'

viruses <- read_tsv(viruses_filename)
```

@@

TO: these rules are not written up correctly

VIBRANT
    Quality == "High Quality Draft": +1
    Quality == "Medium Quality Draft": +1
    Quality == "Low Quality Draft" & provirus == TRUE: +0.5

Virsorter2
    Viral >= 50: +0.5
    Viral >= 0.95: +1
    Hallmark > 2: +1

Virsorter
    category ==  1,2,4,5: +1
    category == 3,6: +0.5

DeepVirFinder:
    Score >= 0.7: +0.5

Kaiju:
    Kaiju_viral = "cellular organisms": -1
    Kaiju_viral = "Viruses": +1

CheckV
    viral_genes == 0 and host_genes >= 1: keep_score = 0
    If 3*viral_genes <= host_genes: keep_score = 0
    If length > 50,000 and hallmark == 0: keep_score = 0
    If %unknown >= 75: +0.5

@@

```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                include_tuning_not_viral=FALSE,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))

  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="complete circular"] <- keep_score[input_seqs$vibrant_quality=="complete circular"] + 1
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$checkv_provirus=="Yes"] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$viral>=50] <- keep_score[input_seqs$viral>=50] + 0.5
    keep_score[input_seqs$viral>=95] <- keep_score[input_seqs$viral>=95] + 0.5
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5
  }

  if (include_deepvirfinder) {
    keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.7 & input_seqs$checkv_length<20000] + 0.5
    keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] <- keep_score[input_seqs$score>=0.9 & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
    keep_score[input_seqs$Kaiju_Viral=="Viruses"] <- keep_score[input_seqs$Kaiju_Viral=="Viruses"] + 0.5
    keep_score[input_seqs$hallmark>2] <- keep_score[input_seqs$hallmark>2] + 1
    keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] <- keep_score[input_seqs$percent_unknown>=75 & input_seqs$checkv_length<50000] + 0.5
    keep_score[input_seqs$percent_viral>=50] <- keep_score[input_seqs$percent_viral>=50] + 0.5
  }

  if (include_tuning_not_viral) {
    keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] <- keep_score[input_seqs$Kaiju_Viral=="cellular organisms"] - 0.5
    keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] <- keep_score[input_seqs$checkv_host_genes>50 & input_seqs$checkv_provirus=="No"] - 1
    keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] <- keep_score[input_seqs$checkv_viral_genes==0 & input_seqs$checkv_host_genes>=1] - 1
    keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] <- keep_score[((input_seqs$checkv_viral_genes*3) <= input_seqs$checkv_host_genes) & input_seqs$checkv_provirus=="No"] - 1 # consider accounting for RNA viruses
    keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_length>500000 & input_seqs$hallmark<=1] - 1
    keep_score[input_seqs$checkv_completeness<1 & input_seqs$hallmark<=1] <- keep_score[input_seqs$checkv_completeness>1 & input_seqs$hallmark<=1] - 0.5 
    keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] <- keep_score[input_seqs$checkv_completeness<=75 & input_seqs$checkv_length<=5000] - 0.5 # helped with protist contamination
  }

@@

}
```

@@

```{r}
combos_list <- data.frame(toolcombo=rep(0, 64),
                          tune_not_viral=rep(0, 64),
                          DVF=rep(0, 64),
                          tune_viral=rep(0, 64),
                          VIBRANT=rep(0, 64),
                          VS=rep(0, 64),
                          VS2=rep(0, 64))
p <- 1

for (i in c(0,1)){
  for (j in c(0,1)){
    for (k in c(0,1)){
      for (l in c(0,1)){
        for (m in c(0,1)){
          for (n in c(0,1)){
            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
            combos_list$toolcombo2[p] <- paste(if(i){"tnv"}else{"0"},if(j){"DVF"}else{"0"},
                                               if(k){"tv"}else{"0"},if(l){"VB"}else{"0"},
                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
            combos_list$tune_not_viral[p] <- i
            combos_list$DVF[p] <- j
            combos_list$tune_viral[p] <- k
            combos_list$VIBRANT[p] <- l
            combos_list$VS[p] <- m
            combos_list$VS2[p] <- n
            p <- p+1
          }
        }
      }
    }
  }
}

combos_list <- combos_list[-1,]
```

```{r}
viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
                          num_viruses=rep(0, nrow(combos_list)))

for (i in 1:nrow(combos_list)) {
  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
                                            include_virsorter = combos_list$VS[i],
                                            include_virsorter2 = combos_list$VS2[i],
                                            include_tuning_viral = combos_list$tune_viral[i],
                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
                                            include_deepvirfinder = combos_list$DVF[i])

@@

  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
}

num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
num_viruses$numtools <- as.factor(num_viruses$numtools)
```

```{r}
ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
                                  color=numtools, fill=numtools)) +
  geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, ntv, VB, VS, VS2)") +
  ylab("Num Viruses Predicted")

ggplot(num_viruses, aes(x=toolcombo2, y=num_viruses, 
                                  color=numtools, fill=numtools)) +
  geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tnv, DVF, tv, VB, VS, VS2)") +
  ylab("Num Viruses Predicted")
```

```{r}
ggplot(num_viruses, aes(x=numtools, y=num_viruses)) +
  geom_boxplot(aes(color=numtools)) +
  geom_point(aes(color=numtools, fill=numtools)) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Number of Tools") +
  ylab("Num Viruses Predicted")
```

```{r}
viral_scores_nozeros <- viral_scores[rowSums(viral_scores)>0,]
viral_scores_nozeros <- viral_scores_nozeros + 1
viral_scores_nozeros <- as.data.frame(viral_scores_nozeros)

colnames(viral_scores_nozeros) <- num_viruses$toolcombo2
```

@@

```{r}
tooldata <- num_viruses

rownames(tooldata) <- tooldata$toolcombo2
```

```{r}
physeq_pooled <- phyloseq(otu_table(viral_scores_nozeros, taxa_are_rows = T),
                                     sample_data(tooldata))
```

```{r}
ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw() +
  geom_label(label=tooldata$toolcombo)

phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw()
```
to do: try coloring above based on the F1 scores of the testing set on each combination

```{r}
bray_dist <- phyloseq::distance(physeq_pooled, method="bray")
clusters <- hclust(dist(bray_dist))
plot(clusters)

myclusters <- cutree(clusters, h=0.8)
```

```{r}
names(myclusters[myclusters==1])
names(myclusters[myclusters==2])
names(myclusters[myclusters==3])
names(myclusters[myclusters==4])
names(myclusters[myclusters==5])

myclusters_df <- tibble(combo=names(myclusters),
                            cluster_index=myclusters)

myclusters_df <- separate(myclusters_df, col=combo, into=c("CheckV", "DVF",
                                                            "Kaiju", "VIBRANT",
                                                            "VirSorter", "VirSorter2"),
                          sep=" ", remove = F)

>>>>

---
title: "Building Metagenome Viral Set and Visualization Template"
author: Bridget Hegarty, James Riddell
date: 10-26-2022
output: html_notebook
---
This Rmarkdown file takes outputs from viral identification tools and merges
them into one file, "{dataset}_merged_viral_id_outputs.csv", which is the input
for identifying_viral_metrics_environmental.Rmd.

@@

```{r setup-library}
library(ggplot2)
library(stringr)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)

```{r}
# dataset name for file organization and outputs for metrics file
dataset_name <- "drinking_water_uknl"
#name from building_env_data_viral_features_set.Rmd
viruses_filename <- 'merged_viral_feature_outputs.tsv'
# path of merged dataset
# path = "/Users/riddellj/Documents/Research/VSTE/VSTE_Old_Pull/EnvironmentalData/DrinkingWaterMetagenomes/"
path = "/Users/riddellj/Documents/GitHub/VSTE/EnvironmentalData/DrinkingWaterMetagenomes/"
setwd(path)
viruses <- read_tsv(paste(dataset_name, viruses_filename, sep='_'))
# write_csv(viruses[viruses$assembly %in% c("49639", "49640", "49641", "49642"),]paste(dataset_name, viruses_filename, 'virome', sep='_'))
```

@@

VIBRANT
    Quality == "Complete Circular": +1
    Quality == "High Quality Draft": +1
    Quality == "Medium Quality Draft": +1
    Quality == "Low Quality Draft" & provirus == TRUE: +0.5

Virsorter2
    Viral >= 50: +0.5
    Viral >= 0.95: +0.5
    RNA >= 0.9: +1
    lavidaviridae >= 0.9: +1
    NCLDV >= 0.9: +1

Virsorter
    category ==  1,4: +1
    category == 2,5: +0.5

DeepVirFinder:
    Score >= 0.7: +0.5
    Score >= 0.9: +0.5

Tuning - Not Viral:
    Kaiju_viral = "cellular organisms": -0.5
    If host_genes >50 and NOT checkv_provirus: -1 
    If viral_genes == 0 and host_genes >= 1: -1
    If 3*viral_genes <= host_genes and NOT checkv_provirus: -1
    If length > 50,000 and hallmark <=1: -1
    If length < 5000 and checkv completeness <= 75: -0.5

Tuning - Viral:
    Kaiju_viral = "Viruses": +1
    If %unknown >= 75 and length < 50000: +0.5
    If %viral >= 50: +0.5
    Hallmark > 2: +1

@@

Function that allows for the comparing pieces of the pipeline
```{r getting_viral_set_1}
getting_viral_set_1 <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
                                include_tuning_not_viral=FALSE,
                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T, ntv_5=F,
                                include_virsorter=FALSE) {

  keep_score <- rep(0, nrow(input_seqs))
  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$provirus] + 0.5
  }

  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
  }

  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
  }

  if (include_deepvirfinder) {
     # add if DVF calls viral
    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
  }

  if (include_tuning_viral) {
    keep_score <- tuning_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=tv_1,
                         include_tv_2=tv_2,
                         include_tv_3=tv_3,
                         include_tv_4=tv_4)
  }

  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score <- tuning_not_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=ntv_1,
                         include_tv_2=ntv_2,
                         include_tv_3=ntv_3,
                         include_tv_4=ntv_4,
                         include_tv_5=ntv_5)

  }

@@

}
tuning_viral <- function(current_score, ins, 
                         include_tv_1=T,
                         include_tv_2=T,
                         include_tv_3=T,
                         include_tv_4=T) {
    # tuning addition
  if (include_tv_1) {
    current_score[ins$hallmark>2] <- current_score[ins$hallmark>2] + 1
  }
  if (include_tv_2) {
    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
  }
    #note: had tried pulling out this rule, but better recall without sacrificing precision with it in
    #note: having kaiju removal rule didn't help
  if (include_tv_3) {
    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
  }
  if (include_tv_4) {
    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
  }
  return(current_score)
}
tuning_not_viral <- function(current_score, ins, 
                         include_tv_1=T,
                         include_tv_2=T,
                         include_tv_3=T,
                         include_tv_4=T,
                         include_tv_5=F) {
    # tuning removal
  if (include_tv_2) {
    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
  }
  if (include_tv_3) {
    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$provirus] - 1
  }

  if (include_tv_5) {
    # remove if DVF calls it not viral
    current_score[ins$score<=0.7 & ins$pvalue<=0.05] <- current_score[ins$score<=0.7 & ins$pvalue<=0.05] - 1 
  }

  if (include_tv_1) {
    current_score[(ins$checkv_host_genes>50) & !ins$provirus] <- -3
  }

  if (include_tv_4) {
    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- -3
  }

  return(current_score)
}
```

```{r}
viruses$keep_score_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)
viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
```

```{r}
sum(viruses$keep_score_all>=1)
sum(viruses$keep_score_high_recall>=1)
sum(viruses$keep_score_high_precision>=1)
sum(viruses$keep_score_high_MCC>=1)
sum(viruses$keep_score_all>=0.5)
sum(viruses$keep_score_high_recall>=0.5)
sum(viruses$keep_score_high_precision>=0.5)
sum(viruses$keep_score_high_MCC>=0.5)
```

```{r}
write_csv(viruses[,c(1,43:45)], paste(dataset_name, 'top_viral_scores.csv', sep="_"))
```

not having the rule, gives the same

@@

```{r}
combos_list <- data.frame(toolcombo=rep(0, 64),
                          tune_not_viral=rep(0, 64),
                          DVF=rep(0, 64),
                          tune_viral=rep(0, 64),
                          VIBRANT=rep(0, 64),
                          VS=rep(0, 64),
                          VS2=rep(0, 64))
p <- 1
for (i in c(0,1)){
  for (j in c(0,1)){
    for (k in c(0,1)){
      for (l in c(0,1)){
        for (m in c(0,1)){
          for (n in c(0,1)){
            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
            combos_list$toolcombo2[p] <- paste(if(i){"tv"}else{"0"},if(j){"DVF"}else{"0"},
                                               if(k){"tnv"}else{"0"},if(l){"VB"}else{"0"},
                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
            combos_list$tune_not_viral[p] <- k
            combos_list$DVF[p] <- j
            combos_list$tune_viral[p] <- i
            combos_list$VIBRANT[p] <- l
            combos_list$VS[p] <- m
            combos_list$VS2[p] <- n
            p <- p+1
          }
        }
      }
    }
  }
}
combos_list <- combos_list[-1,]
```

```{r}
viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
                          num_viruses=rep(0, nrow(combos_list)))
for (i in 1:nrow(combos_list)) {
  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
                                            include_virsorter = combos_list$VS[i],
                                            include_virsorter2 = combos_list$VS2[i],
                                            include_tuning_viral = combos_list$tune_viral[i],
                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
                                            include_deepvirfinder = combos_list$DVF[i])

@@

  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
}
num_viruses$numtools <- stringr::str_count(num_viruses$toolcombo, "1")
num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
num_viruses$numtools <- as.factor(num_viruses$numtools)
```

```{r}
png(paste(dataset_name, 'virus_predicted.png', sep='_'))
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
# add this
ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
                                  color=numtools, fill=numtools)) +
  geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Num Viruses Predicted") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1))
dev.off()
```

```{r}
viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
num_assemblies <- dim(table(viruses$assembly))
assemblies <- names(table(viruses$assembly)) # add assemblies if not in viruses yet
num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)*num_assemblies),
                          toolcombo2=rep(0, nrow(combos_list)*num_assemblies),
                          assembly=rep(0,nrow(combos_list)*num_assemblies),
                          num_viruses=rep(0, nrow(combos_list)*num_assemblies),
                          prop_vir=rep(0, nrow(combos_list)*num_assemblies)
                          )
n <- 1
for (i in 1:nrow(combos_list)) {
  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
                                            include_virsorter = combos_list$VS[i],
                                            include_virsorter2 = combos_list$VS2[i],
                                            include_tuning_viral = combos_list$tune_viral[i],
                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
                                            include_deepvirfinder = combos_list$DVF[i])
  for (j in 1:num_assemblies) {
    if (max(viral_scores[,i])<=0) {
      num_viruses$num_viruses[n] <- 0
    }
    else {
      num_viruses$num_viruses[n] <- sum(viral_scores[viruses$assembly==assemblies[j],i]>=1)
      num_viruses$prop_vir[n] <- num_viruses$num_viruses[n]/table(viruses$assembly)[[j]]
    }

    num_viruses$toolcombo[n] <- combos_list$toolcombo[i]
    num_viruses$toolcombo2[n] <- combos_list$toolcombo2[i]
    num_viruses$assembly[n] <- assemblies[j]

    n <- n+1
  }
}
num_viruses$numtools <- stringr::str_count(num_viruses$toolcombo, "1")
num_viruses <- num_viruses[order(num_viruses$prop_vir, decreasing=F),]
num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
num_viruses$numtools <- as.factor(num_viruses$numtools)
```

```{r}
write_csv(num_viruses, paste(dataset_name, "num_viruses.csv", sep="_"))
```

```{r}
png(paste(dataset_name, 'prop_viral_assembly.png', sep='_'))
ggplot(num_viruses, aes(x=toolcombo, y=prop_vir, 
                                  shape=numtools, color=assembly, fill=assembly)) +
  geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral")
dev.off()
```

only plotting the average

```{r}
num_viruses$numtools <- as.numeric(num_viruses$numtools)
median_prop_vir_by_tool <- num_viruses %>% 
  select(toolcombo, prop_vir, numtools) %>%
  group_by(toolcombo) %>%
  summarize(median=median(prop_vir),
         #sd=sd(prop_vir),
         numtools=mean(numtools)) 
median_prop_vir_by_tool$toolcombo <- factor(median_prop_vir_by_tool$toolcombo, 
                                          levels = median_prop_vir_by_tool$toolcombo[order(median_prop_vir_by_tool$median, decreasing = F)])
median_prop_vir_by_tool$numtools <- as.factor(median_prop_vir_by_tool$numtools)
```

```{r}
png(paste(dataset_name, 'prop_viral_average.png', sep='_'))
ggplot(median_prop_vir_by_tool, aes(y=median, x=toolcombo,
                                  color=numtools, fill=numtools, shape=numtools))  +
  geom_hline(yintercept=0, color="grey") +
  geom_point() +
#  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
#                 position=position_dodge(.9)) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1))
dev.off()
```
```{r}
png(paste(dataset_name, 'boxplot.png', sep='_'))
ggplot(median_prop_vir_by_tool, aes(x=numtools, y=median, 
                                  color=numtools, fill=numtools)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Proportion Viruses (%)") +
  xlab("Number of Tools") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1))
dev.off()
```

by only a subset of the assemblies (UK/NL)

```{r}
num_viruses_uknl <- num_viruses
num_viruses_uknl$residual <- "residual"
num_viruses_uknl$residual[num_viruses_uknl$assembly=="Amsterdam_Leid" |
                          num_viruses_uknl$assembly=="Amsterdam_Wee" |
                          num_viruses_uknl$assembly=="Nieuwkoop_DHB" |
                          num_viruses_uknl$assembly=="South_Holland_Dla" |
                          num_viruses_uknl$assembly=="Spannenburg"
                          ] <- "no residual"
```

```{r}
num_viruses_uknl$numtools <- as.numeric(num_viruses_uknl$numtools)
mean_prop_vir_by_tool <- num_viruses_uknl %>% 
  select(toolcombo, prop_vir, numtools, residual) %>%
  group_by(toolcombo, residual) %>%
  summarize(prop_vir=mean(prop_vir),
         #sd_per_vir=sd(prop_vir),
         numtools=mean(numtools)) 
mean_uknl <- mean_prop_vir_by_tool %>% 
  group_by(toolcombo) %>%
  summarize(mean=mean(prop_vir)) 
mean_prop_vir_by_tool$toolcombo <- factor(mean_prop_vir_by_tool$toolcombo, 
                                          levels = mean_uknl$toolcombo[order(mean_uknl$mean, decreasing = F)])
mean_prop_vir_by_tool$numtools <- as.factor(mean_prop_vir_by_tool$numtools)
```

```{r}
num_viruses_uknl$toolcombo <- factor(num_viruses_uknl$toolcombo, 
                                          levels = levels(mean_prop_vir_by_tool$toolcombo))
num_viruses_uknl$numtools <- as.factor(num_viruses_uknl$numtools)
```

```{r}
png(paste(dataset_name, 'prop_viral_residual.png', sep='_'))
ggplot(num_viruses_uknl) +
  geom_hline(yintercept=0, color="grey") +
  geom_point(aes(color=numtools, fill=numtools, y=prop_vir, x=toolcombo, shape=residual), alpha=0.5) +
  geom_point(data=mean_prop_vir_by_tool, aes(x=toolcombo, 
                                               y=prop_vir,
                                             shape=residual),
             color="black") +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1))
dev.off()
```
```{r}
#t.test(num_viruses_uknl$prop_vir[num_viruses_uknl$residual=="residual"],
#       num_viruses_uknl$prop_vir[num_viruses_uknl$residual=="no residual"]
#       )
```

ordination: sample by tool combo

@@

make the sample by tool combo matrix
```{r}
num_viruses_uknl_wide <- num_viruses_uknl %>% 
  select(toolcombo, assembly, num_viruses) %>%
  pivot_wider(names_from=assembly, values_from=num_viruses)
num_viruses_uknl_wide <- as.data.frame(num_viruses_uknl_wide)
rownames(num_viruses_uknl_wide) <- num_viruses_uknl_wide$toolcombo
num_viruses_uknl_wide <- num_viruses_uknl_wide[,-1]
num_viruses_uknl_wide <- num_viruses_uknl_wide[rowSums(num_viruses_uknl_wide)>0,]
```

```{r}
tooldata <- combos_list
rownames(tooldata) <- tooldata$toolcombo
tooldata <- tooldata[c(rownames(num_viruses_uknl_wide)),]
tooldata$numtools <- str_count(tooldata$toolcombo, "1")
tooldata$numtools <- as.factor(tooldata$numtools)
tooldata$total_viruses <- rowSums(num_viruses_uknl_wide)
```

to reduce arch effect:
```{r}
num_viruses_uknl_wide <- num_viruses_uknl_wide/rowSums(num_viruses_uknl_wide)
```

```{r}
physeq_pooled <- phyloseq::phyloseq(phyloseq::otu_table(num_viruses_uknl_wide, taxa_are_rows = F),
                                     phyloseq::sample_data(tooldata))
```

```{r}
ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
```
